---
title: "YueZhang_A02_CodingBasics.Rmd"
author: "Yue Zhang"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1. 
seq1 <- seq(1,100,4) #from, to, by
seq1

#2. 
meanseq1 <- mean(seq1)
meanseq1

medianseq1 <- median(seq1)
medianseq1

#3.
meanseq1 < medianseq1

```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}
#5.
a <- c('yue','jilin','will','fangyi','zoe') #character
a
b <- c(90,93,89,94,49) #numeric
b
c <- c(TRUE,TRUE,TRUE,TRUE,FALSE) #logical
c

#6.
class(a)
class(b)
class(c)

#7.
df_studenttest <- data.frame(a,b,c)
df_studenttest

#8.
names(df_studenttest) <- c('name','test_score','pass')
df_studenttest
```

9.  QUESTION: How is this data frame different from a matrix?

> Answer:
Matrices can only contain a single class of data, while data frames can consist of many different classes of data.

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r}
#10.
test_score <- function(x){
  ifelse(x >= 50, TRUE, FALSE)
}

#11.
test_score(b)

```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer:
'ifelse'worked. 'if' and 'else' function only deal with a single value. So if we put a vector in the 'if' statement, it will only check the very first element and issue a warning. But the 'ifelse' can check the condition for every element of a vector and select elements from the specified vector depending upon the result.
